# Author Statistics {#sec:authors}


\question{Distribution of "paper readiness" and relationship to author experience}{paper-readiness}

  - http://www.ieee-security.org/TC/Reports/2017/SP2017-PCChairReport.pdf
  - https://www.nsf.gov/statistics/2018/nsb20181/assets/968/tables/tt05-18.pdf

\question{Does submission order affect paper's chances?}{sub-order}

  - http://onlinelibrary.wiley.com/doi/10.1002/asi.22747/abstract
  
```{r sub-order, echo=F, warning=F}
#papers_grades_order <- survey %>% 
#  group_by(paper_id) %>% 
#  summarise(gr_overall = mean(gr_overall)) %>%
#  left_join(select(papers, c("key", "months_to_gs", "months_to_eprint")), by=c("paper_id" = "key"))  %>%
#  drop_na(gr_overall, months_to_gs, months_to_eprint)

#cor.test(papers_grades_order$gr_overall,  papers_grades_order$months_to_gs)
#cor.test(papers_grades_order$gr_overall, papers_grades_order$months_to_eprint)
```

\question{Compare "overall" average to confidence-weighted "overall" average}{confidence}

  - [Search references]
  
```{r confidence, echo=F, warning=F}
conf_tbl <- survey %>%
  drop_na(gr_overall, gr_confidence) %>%
  group_by(paper_id) %>%
  mutate(conf_weight = gr_overall * gr_confidence) %>%
  summarise(gr_overall_avg = mean(gr_overall), gr_conf_avg = mean(conf_weight))

cor.test(conf_tbl$gr_overall_avg, conf_tbl$gr_conf_avg) #strong correlation between overall and confidence weighted overall (obviously)
#now, make nice chart for it
library(reshape)
melted <- melt(as.data.frame(conf_tbl), id=c("paper_id"))
ggplot(melted, aes(x=variable, y=value, color=variable)) +
  geom_boxplot() + 
  stat_summary(fun.y=mean, geom="point", shape=23, size=4)  +
  xlab(NULL) + 
  ylab("Grade") +
  theme_minimal()

t.test(conf_tbl$gr_conf_avg, conf_tbl$gr_overall_avg)

#blurb about more confident if positive about data
#cavat that there is no rejected paper reviews to compare to
```

\question{Produce initial stats like these}{stats}

  - https://publons.com/community/gspr

\question{Explain zero correlation between prior_subs and gr_overall}{prior-overall}

  - [Search references] + diversity/survey paper


\question{How to deal with publication bias?}{pub-bias}

  - https://www.nature.com/articles/s41562-016-0034
  - http://www.nature.com/news/let-s-think-about-cognitive-bias-1.18520
  - http://www.nature.com/news/how-scientists-fool-themselves-and-how-they-can-stop-1.18517
  - http://www.nature.com/news/tool-for-detecting-publication-bias-goes-under-spotlight-1.21728
  - http://www.nature.com/news/replication-studies-bad-copy-1.10634



----

