---
title: Why So Few Women in Computer Systems Research?
author:
  - name: Alexis Richter
    affiliation: 1
    corresponding: false
    email: ajrichter@reed.edu
  - name: Josh Yamamoto
    affiliation: 1
    corresponding: false
    email: yamamojo@reed.edu
  - name: Eitan Frachtenberg
    affiliation: 1
    corresponding: true
    email: eitan@reed.eud
affiliation:
    address: Department of Computer Science, Reed College

bibliography: ../sysconf.bib
output:
  bookdown::pdf_book:
    template: main.tex
    keep_tex: yes
    citation_package: natbib
abstract: |
  Numerous explanations have been proposed for the underrepresentation of women in computer science (CS) research. Using an expansive dataset on CS conference authors, we ask which of these explanations apply even more to researchers in computer systems, where the representation of women is even lower.
---

```{r code = readLines("../load_data.R"), echo = F, message = F}
```


```{r setup, echo=F, message=F, warning=F, cache=F}
library('tidyverse')
library('rjson')
library('kableExtra')

# Colors:
cwomen <- "#7704FF"
cmen <- "#00C3AA"
cmen <- "gray70"
cwomen <- "gray30"
cavg <- "#FF10F0"
cavg <- "#CB00E6"


# This function reads in a .json conference file and for each paper in the conference it extracts the author names, paper id,
# field, and conference id into a data frame. It then normalizes the names using 'normalized_author_name' and moves
# on to the next paper, stacking the new data frame with those that have already been created each time.
# Finally it normalizes the author names and joins with a gender mapping df to get the corresponding genders.

json_to_df <- function(conf_name, gender_mapping) {
  json_fpath <- paste0(toplevel, "data/conf/", conf_name, ".json")
  conf <- rjson::fromJSON(file = json_fpath)
  confname <- conf[[1]]
  field <- conf[['field']]
  subfield <- conf[['subfield']]
  papers <- conf[['papers']]
  submissions <- conf[['submissions']]

  conf_info_df <- data.frame()

  for (i in seq_along(papers)) {
    paper_id <- unlist(papers[[i]][1])
    authors <- as.data.frame(papers[[i]][3]) %>%
      mutate(conf = confname,
             field = field,
             subfield = subfield,
             paper_id = paper_id,
             n_submissions = submissions)

    conf_info_df <- bind_rows(conf_info_df, authors)
  }

  conf_info_df <- conf_info_df %>%
    mutate(authors = map_chr(authors, normalized_author_name)) %>%
    mutate(authors = str_replace(authors, "\\s+", " ")) %>%
    mutate(paper_id = as.factor(paper_id)) %>%
    rename(name = authors)

  conf_info_df %>%
    left_join(gender_mapping, by = "name")
}

######################
# Here we create a vector of the conference names and use the 'map_dfr' function to apply 'json_to_df' to each conference.
nonsys_confs <-
  c("AAAI", "ACL", "CHI", "CVPR", "FSE", "ICML", "ICSE", "MM", "NIPS", "POPL", "SIGCSE", "SIGGRAPH", "SODA", "STOC", "WSDM", "WWW", "ITICSE", "FOCS", "TACAS")

nonsys_authors <- nonsys_confs %>%
  map_dfr(~json_to_df(.x, all_genders)) %>%
  mutate(across(c(field, subfield, gender, paper_id), factor)) %>%
  mutate(across(.cols = everything(), na_if, NA))


###########
name_combination_creator <- function(df)
{
  authors <- df$name
  if (length(authors) <= 1) {
    return (data.frame(name1 = c(), name2 = c()))
  }

  ret <- as.data.frame(t(combn(authors, 2)))
  names(ret) <- c("name1", "name2")
  ret <- rbind(ret, data.frame(name1 = ret$name2, name2 = ret$name1))
  return(ret)
}

#######
# All-to-all coauthor pairs for non-sys and for sys:
nonsys_collaboration_pairs <- nonsys_authors %>%
  group_by(paper_id) %>%
  group_modify(~name_combination_creator(.x)) %>%
  ungroup() %>%
  left_join(all_genders, by = c("name2" = "name")) %>%
  rename(gender2 = gender) %>%
  left_join(all_genders, by = c("name1" = "name")) %>%
  rename(gender1 = gender) %>%
  dplyr::select(paper_id, name1, gender1, name2, gender2)

sys_collaboration_pairs <- coauthors %>%
  left_join(persons, by = c("name1" = "name", "gs_email1" = "gs_email")) %>%
  rename(gender1 = gender, paper_id = paper) %>%
  left_join(persons, by = c("name2" = "name", "gs_email2" = "gs_email")) %>%
  rename(gender2 = gender) %>%
  dplyr::select(paper_id, name1, gender1, name2, gender2) %>%
  distinct(name1, name2, paper_id, .keep_all = T)

all_pairs <- rbind(nonsys_collaboration_pairs, sys_collaboration_pairs)

#####################
gender_summary_table <- function(data)
{
  data %>%
    summarise(.groups = "keep",
      n_papers = n_distinct(paper_id, conf),
      n_na = sum(is.na(gender)),
      n_male = sum(gender == "M", na.rm = T),
      n_female = sum(gender == "F", na.rm = T),
      prop_female = n_female / (n_female + n_male)
    ) %>%
    arrange(desc(prop_female)) %>%
    knitr::kable()
}

#######

sys_auths <- roles %>%
  filter(role == "author") %>%
  left_join(persons, by = c("name" = "name", "gs_email" = "gs_email")) %>%
  dplyr::select(name, gender, key, gs_email) %>%
  mutate(conf = as.factor(gsub("_\\d\\d\\d$", "", key))) %>%
  left_join(all_confs, by = c("conf" = "conference")) %>%
  rename(key = key.x) %>%
  dplyr::select(name, gender, key, field, subfield)

nonsys_auths <- nonsys_authors %>%
  dplyr::select(-conf, -n_submissions) %>%
  dplyr::select(name, gender, paper_id, field, subfield) %>%
  rename(key = paper_id) %>%
  mutate(gs_email = NA)

full_auths <- bind_rows(sys_auths, nonsys_auths) %>%
  group_by(key) %>%
  mutate(coauthors = n() - 1) %>%
  mutate(last_gender = last(gender), is_last = name == last(name)) %>%
  ungroup() %>%
  group_by(name, gender) %>%
  mutate(prod = n(), fractional_prod = sum(1 / (coauthors + 1))) %>%
  ungroup()

full_auths$subfield = recode_factor(full_auths$subfield,
                                    "Heterogeneous Computing" = "Computer Architecture",
                                    "Network" = "Communications",
                                    "Virtualization" = "Operating Systems",
                                    "Energy" = "Computer Architecture",
                                    "Architecture" = "Computer Architecture")

full_auths$field = recode_factor(full_auths$field,
                                 "Computer Science Education" = "Computer science education",
                                 "Human-Computer Interaction" = "Human-computer interaction",
                                 "Knowledge Systems" = "Knowledge systems",
                                 "Software Engineering & Languages" = "Software engineering and languages",
                                 "Artificial Intelligence" = "Artificial intelligence",
                                 "Systems" = "Computer systems",
                                 "Theory and Algorithms" = "Theory and algorithms"
)

gen_auths <- full_auths %>% drop_na(gender)

########
pairs_with_field <- full_auths %>%    # Add field and subfield to all_pairs
  mutate(paper_id = key) %>%
  group_by(paper_id) %>%
  summarize(.groups = "keep", field = first(field), subfield = first(subfield)) %>%
  right_join(all_pairs) %>%
  ungroup() %>%
  drop_na(gender1) %>%
  drop_na(gender2) %>%
  group_by(subfield) %>%
  mutate(internal = sum(gender1 == gender2), external = sum(gender1 != gender2)) %>%
  mutate(ei_index = (external - internal) / (external + internal)) %>%
  mutate(w_prob = sum(gender1 == "F" & gender2 == "F") / sum(gender1 == "F"))
  
#######
productivity_diff <- function(sf)
{
  men <- filter(gen_auths, gender == "M", subfield == sf)
  women <- filter(gen_auths, gender == "F", subfield == sf)
  paste0(round(mean(men$prod) - mean(women$prod), 2), "; ",
         report_test(t.test(men$prod, women$prod)))
}

#### Summary table of properties of subfields.
tmp <- pairs_with_field %>%
  group_by(subfield, field) %>%
  summarize(.groups = "keep", ei_index = first(ei_index), wprob = first(w_prob))

field_summary <- gen_auths %>%
  group_by(field) %>%
  summarize(.groups = "keep",
            Authors = n(),
            Coauthors = n() / n_distinct(key),
            FAR = sum(gender == "F") / n()) %>%
  ungroup()

lf <- filter(full_auths, last_gender == "F")  # All authors in papers where the last author (senior) is female
lm <- filter(full_auths, last_gender == "M")  # All authors in papers where the last author (senior) is male
tot_fm <- nrow(filter(lm, gender == "F", !is_last))  # Total no. of nonsenior females in papers with senior male
tot_ff <- nrow(filter(lf, gender == "F", !is_last))  # Total no. of nonsenior females in papers with senior female
tot_mf <- nrow(filter(lf, gender == "M", !is_last))  # Total no. of nonsenior males in papers with senior female
tot_mm <- nrow(filter(lm, gender == "M", !is_last))  # Total no. of nonsenior males in papers with senior male

# Repeat these calculations but for systems papers only:
sys_lf <- filter(lf, key %in% sys_auths$key)
sys_lm <- filter(lm, key %in% sys_auths$key)
nonsys_lf <- filter(lf, !(key %in% sys_auths$key))
nonsys_lm <- filter(lm, !(key %in% sys_auths$key))
```


# Introduction {-}


```{r field-aggregation, echo = F, cache = T}
data.frame(
    Field = field_summary$field,
    "Total authors" = field_summary$Authors,
    "Female authors" = paste0(round(100 * field_summary$FAR, 1), "%"),
    check.names = F
) %>%
  bind_rows(data.frame(check.names = F,
    "Field" = "Overall",
    "Total authors" = sum(field_summary$Authors),
    "Female authors" = paste0(round(100 * sum(gen_auths$gender == "F") / nrow(gen_auths), 1.0), "%"))) %>%
    knitr::kable(format = "latex",
               booktabs = T,
               align = "lrr",
               linesep = c(rep("", 6), "\\addlinespace"),
               caption = "Percentage of female authors in peer-reviewed conference papers across fields of computer science, omitting authors with ambiguous gender \\cite{yamamoto22:collab}. Authors count multiple times when they appear on more than one paper. The difference in female author ratio between systems and nonsystems fields is statistically significant ($p<10^{-9}$).")
#  column_spec(2, width = "2.5cm")
#  kable_styling(font_size = 7)

```


<!-------------------------------------------------------------------------------------------------------------->




In a recent study, we found that the ratio of female-to-male authors in computer systems conferences is particularly low, even compared to the rest of CS [@yamamoto22:collab].
The large and statistically significant underrepresentation cannot be fully explained by review bias, differences in collaboration patterns, and numerous other demographic and conference factors alone [@frachtenberg22:gender-gap].

Having previously answered the question of "how many women in computer systems research?", we now ask "why so few?", expanding beyond measurable facts and venturing into the domain of hypothesis and speculation.
This simple question likely has no simple answers, and this article does not attempt to provide any definitive ones either.
Instead, we attempt to break down this big question into a number of smaller, speculative questions, and gather what evidence we can find to provide, if not outright answers, at least justification for the questions and for future investigation.

There are many dimensions---economic, societal, historical, and structural---that influence this gender gap [@avolio20:factors; @singh13:stemming].
We may never be able to explain it fully; however, we can examine many of the factors that have been previously identified as possible contributors to the gender gap in CS and other sciences, in an attempt to assess whether they have an outsize effect on systems in particular.^[For this study, we define systems as the study and engineering of concrete computing systems, which includes research topics such as operating systems, computer architectures, data storage and management, compilers, parallel and distributed computing, and computer networks.]
Based on our experience in the field, our conversations with fellow female researchers about their experiences, and the existing literature on the gender gap in other fields, we attempt to examine possible contributors to the gender gap in CS and other sciences in the specific context of the systems field.

In this viewpoint, we examine ten such factors, supporting our discussion with quantitative evidence when possible.
Our main source of data is a manually curated dataset of `r nrow(sys_confs) + length(nonsys_confs)` reputable CS conferences, all from 2017.
The conferences from the fields we identified as *systems* included a total of `r fmt(nrow(sys_authors))` unique authors, with the remaining `r fmt(nrow(nonsys_authors))` unique authors from non-systems conferences.
All `r fmt(n_distinct(full_auths$key))` peer-reviewed research-track papers in these conferences and their authors were included in the data.
We manually matched most authors with a *perceived gender*, based on pronouns and photos, rather than using automated approaches using name-based inference, because the accuracy and recall of this inference can be quite low for some populations.
In total, we analyzed `r fmt(length(full_auths$name))` authors (`r fmt(length(unique(full_auths$name)))` unique).
The complete methodological aspects of identifying and cleaning the appropriate data, as well as the main statistical findings about the underrepresentation of women in the field, can be found in our original study [@yamamoto22:collab].

Here, we combine our current dataset with prior data sources to try to evaluate the impact of each of the ten potential factors on the increased gender gap in systems, each presented as a hypothesis that it has a more significant effect on the gender gap in systems in particular.
The ten hypotheses are organized into two groups: six where our data provide evidence for or against an increased effect, and four for which we have no specific data but can speculate based on the characteristics of systems research.
For each factor, we briefly explain its observed effects on the gender gap from previous studies, and then try to find support in the data (or lack thereof) for an increased effect of this factor in systems.

# Factors for which we have quantitative evidence {-}

## H1: Too few women peers and mentors {-}

```{r compute_last_authors, echo=F}
lf <- filter(full_auths, last_gender == "F")  # All authors in papers where the last author (senior) is female
lm <- filter(full_auths, last_gender == "M")  # All authors in papers where the last author (senior) is male
tot_fm <- nrow(filter(lm, gender == "F", !is_last))  # Total no. of nonsenior females in papers with senior male
tot_ff <- nrow(filter(lf, gender == "F", !is_last))  # Total no. of nonsenior females in papers with senior female
tot_mf <- nrow(filter(lf, gender == "M", !is_last))  # Total no. of nonsenior males in papers with senior female
tot_mm <- nrow(filter(lm, gender == "M", !is_last))  # Total no. of nonsenior males in papers with senior male

# Repeat these calculations but for systems papers only:
sys_lf <- filter(lf, key %in% sys_auths$key)
sys_lm <- filter(lm, key %in% sys_auths$key)
nonsys_lf <- filter(lf, !(key %in% sys_auths$key))
nonsys_lm <- filter(lm, !(key %in% sys_auths$key))
```

Paradoxically, a large gender gap in a field could itself be partially caused by an existing large gender gap.
The lack of female mentors, peers, and role models dissuades women from entering or remaining in a field where they can feel unsupported and singled-out [@cheryan17:STEM; @cohoon06:critical; @drury11:female; @frantzana19:women; @kahn17:women].
Additionally, it appears that women in systems have a small but significant preference to publishing with other women---which are already scarce  [@yamamoto22:collab].
This gender homophily is certainly not unique to this field, but it reinforces the notion that having fewer women in systems makes it less inviting to new women.

Moreover, in the overall `r fmt(n_distinct(full_auths$key))` CS papers we looked at, when the senior (last) author is a woman, non-senior authors include
`r pct(tot_ff, nrow(lf))`% women, compared to only
`r pct(tot_fm, nrow(lm))`% when the senior author is a man.
However, of the `r fmt(nrow(papers))` papers in systems, only
`r pct(nrow(sys_lf), nrow(sys_lf) + nrow(sys_lm))`%
have a female senior author, compared to
`r pct(nrow(nonsys_lf), nrow(nonsys_lf) + nrow(nonsys_lm))`%
in the non-systems papers.
The paucity of systems papers with senior women suggests that junior women have about 50% better chance to find a senior woman to collaborate with outside of systems.
 

## H2: Highly collaborative fields are less welcoming to gender minorities {-}

Another potential consequence of having fewer peers and fewer publications for women is that finding similar collaborators for a research project may be more difficult in certain fields [@abbate12:recoding; @zeng16:differences], especially those that typically require larger research teams [@yamamoto22:collab]. So the largely male teams may feel less welcoming to women authors, compared to fields that rely less on large collaborative efforts.
Indeed, in our dataset, systems had the largest team sizes
(`r round(mean(filter(gen_auths, field == "Computer systems")$coauthors), 2)`
mean coauthors per paper vs.
`r round(mean(filter(gen_auths, field != "Computer systems")$coauthors), 2)`
in other fields), possibly because this field often requires large collaborative efforts to implement complex systems.
(Men also averaged 2% larger team sizes than women, but this difference was not statistically significant.)


## H3: Women are better represented in human-centered fields {-}

It has been shown that owing to traditional social roles, women tend to be more people-centered and communally oriented than men in their career choices [@beyer04:deterrents; @diekman15:new; @west13:role].
On the other hand, most science and technology careers are not perceived as focusing heavily on these values [@frantzana19:women; @kahn17:women].
Even within CS, the representation of women appears much higher in fields ostensibly centered on humans, such as human-computer interaction and computer education (Table \@ref(tab:field-aggregation)).
Although many computer systems directly and indirectly benefit society, the field could be perceived as one that primarily focuses on machines, and not on societal or community benefits.
That perception alone may further deter women’s entry into the field.


## H4: Women are underrepresented in engineering fields {-}

CS is a varied discipline that combines aspects from mathematics, algorithmics, and engineering.
In particular, work in systems is very similar to other engineering disciplines, with its emphasis on efficiency, trade-off management, and practical applications.
At least `r pct(218, nrow(papers), 0)`% of the systems papers in our dataset have one or more authors affiliated with an academic engineering department, although the percentage of authors with an electrical or computer engineering degree is potentially even higher.
It is therefore not surprising that the representation of women we find in systems papers is very similar to some estimates for engineers in general [@fox06:engineering; @mattis07:upstream].

<!-- Percentage computed with the following command-line:
pdfgrep -i -n engineering *.pdf | grep ':1:' | egrep -i 'department|school' | awk -F: ' {print $1;}' | sort | uniq | wc
-->


## H5: Fewer women pursue research in industry {-}

Women engineers also appear to be significantly underrepresented in industry research [@ghiasi15:compliance; @tao16:employment].
As a field related to engineering, we might expect systems to exhibit similarly large gender gaps for industry researchers.
In addition, since systems research and development is increasingly central to the success of both technology startups and large-scale companies such as Google, Amazon, and Microsoft, we might also expect the high demand for skilled workers in industry to amplify this factor.

Despite our expectations, the hypothesis that an industry affiliation leads to a higher gender gap in research is only weakly supported by our data.
By extracting the email address associated with each author in our systems papers, we can roughly categorize each author with a uniquely identifiable affiliation as either in government, academia, or industry [@frachtenberg22:gender-gap].
Of the 
`r as <- filter(authors, !is.na(sector), !is.na(gender)); pct(nrow(filter(as, sector == "COM")), nrow(as), 1)`%
authors with an industry affiliation, 
`r pct(nrow(filter(as, sector == "COM", gender == "F")), nrow(filter(as, sector == "COM")), 1)`%
were women, not much lower than the overall rate of women.
<!--(`r report_test(chisq.test(table(as$gender, as$sector == "COM")), show_stat = F)`).-->


## H6: Women leave research careers at a higher rate than men {-}

It is well-known that women in the sciences and CS leave academia at a higher rate than men (the so-called “leaky pipeline”) [ @kahn17:women; @singh13:stemming; @fox06:engineering; @camp02:incredible].
Looking at the ratio of researchers in our dataset that appear to have stopped publishing by 2019, we find, surprisingly, that women did *not* leave research at a higher rate than their male counterparts; only 16.9% of women ostensibly stopped publishing, compared to 24.4% of men.
As another measure of the retention rate of women to senior positions, we can also look at the ratio of female last authors, since in CS and especially in systems, authors are often ordered by contribution, with the senior author at the last position.
In our dataset, only 12.3% of systems papers with three or more authors have ordered them alphabetically.
Women last authors appear in
`r pct(nrow(sys_lf), nrow(sys_lf) + nrow(sys_lm))`% of our systems papers,
which is actually closer to the ratio of all female authors in systems,
`r pct (nrow(filter(gen_auths, gender=="F", field == "Computer systems")), nrow(filter(gen_auths, field == "Computer systems")))`%, compared to the ratios in non-systems papers
(`r pct(nrow(nonsys_lf), nrow(nonsys_lf) + nrow(nonsys_lm))`% and
`r pct (nrow(filter(gen_auths, gender=="F", field != "Computer systems")), nrow(filter(gen_auths, field != "Computer systems")))`% respectively), suggesting a similar or lower attrition rate for women in systems.
Rather than having an attrition problem, it is possible that the field has a bigger challenge attracting women to the field in the first place.

# Factors for which we have no quantitative evidence {-}


## H7: Systems research requires more pre-college programming experience {-}

The gender disparity in pre-college computing and programming experience has been studied extensively as a factor affecting post-secondary participation in CS [@cheryan17:STEM; @cohoon06:critical].
Systems research in particular may require an intimate understanding of programming systems and may benefit from pre-college programming experience, as has been found for the systems subfield of high-performance computing [@frantzana19:women].
Pre-college girls were also found to be less likely than boys to dissect a computer [@beyer04:deterrents].
Since women are less likely to learn about hardware and software at a young age, they may be less experienced than men with the low-level programming tasks required in systems.

## H8: Women are underrepresented in more expensive fields {-}

Studies have shown that in fields where research financial expenditures are higher, women consistently publish less than men [@duch12:possible].
Systems research often requires expensive equipment such as experimental computer architectures, supercomputers, or cloud resources.
It is therefore plausible that the higher cost of systems research leads to lower participation and publication rates by women, as opposed to other, less costly CS fields.

## H9: Systems research implies worse work-life balance {-}

If systems research is indeed more competitive than some other academic fields (for which we found no published evidence), it may require a bigger commitment to work for success.
Similarly, many of the technology companies that attract skilled systems researchers have a reputation for long working hours.
These factors could lead to a work-life balance that is worse for systems than it is for some other CS research fields.
Some women may already feel that a CS career is to reconcile with raising a family [@beyer04:deterrents], and the perception of systems careers as particularly demanding may exacerbate this problem.

## H10: Women's participation rate in a field varies by geography  {-}

Several studies have found that regional, cultural, and societal factors can significantly affect the number of women in a field [@cheryan17:STEM].
For example, even the geographical variance in implicit gender stereotypes can explain gender disparity in math and science performance [@hill10:why].
In our dataset, the representation of women varies considerably by country [@frachtenberg22:gender-gap]. Even among the top-20 countries (by number of authors), representation varies from 2% in Portugal and Japan to 14% in Sweden (11.4% in the U.S, the largest country with 55% of systems authors).
We don't have evidence for gender stereotypes or cultural differences that are specific to computer systems, but we could hypothesize that when it comes to systems, some of the preceding hypotheses have the potential to sharpen preexisting cultural differences towards CS and sciences in general, such as the effects on work/life balance.

---


<!-- With all the evidence pointing out the lack of support, mentorship, and peers for female systems researchers, it is little wonder that the first support group for women in computing, "systers", was created specifically in the systems field [@abbate12:recoding]. -->
The possible explanations for the gender gap we listed here are by all means not exhaustive.
Other well-known factors we didn’t discuss include occupational stereotypes, ability, confidence, demographics, discrimination, bias, competitive work, and a sense of belonging or of being valued.
We found no specific reasons to believe these additional factors have an increased effect in systems research.
However, we have heard anecdotal complaints from women practitioners about the toxic masculine culture in systems labs and some systems conferences^[Read for example Margaret Martonosi's statement on diversity in at the MICRO-50 conference in the ACM SIGARCH blog (https://www.sigarch.org/statement-on-diversity-at-micro-50/).].

Attempting to generalize the ten specific factors we discussed is challenging because the observed aggregate gender gap is the result of many individual decisions based on individuals' circumstances and experiences.
One common thread, however, is the chicken-and-egg nature of the problem.
Most of the factors (causes) are themselves exacerbated by the current dearth of women in the field (effect), which has persisted for years.
We therefore believe that addressing this chicken-and-egg problem cannot succeed with diversity initiatives in hiring or conferences alone [@frachtenberg22:gender-gap].
Such initiatives are clearly welcome, as well as other interventions to reducing the gender gap in CS [@cohoon06:critical; @main17:underrepresentation].
But for systems in particular, we think a concerted effort is warranted to significantly increase the number of women across the field, making it more welcoming to women and minorities in general and bootstrapping a long-term solution.

At an institutional level, one approach may be to hire a cluster of women researchers together instead of piecemeal.
Given the existing gender gap we described, this is admittedly easier said than done; but hiring one woman at a time may actually contribute to the problem rather than the solution, as we described.
To address this challenge, an institute may take a long-term perspective of encouraging a large group of women to enter the field early in graduate school (or even undergraduate), offering them a clear career horizon with reasonable prospects of returning as faculty.
At a national level, governments and large professional organizations such as the ACM could commit to long-term scholarships and grants for women in the field, starting perhaps as early as high school and continuing to support them through the formation of a research lab, perhaps all the way to tenure.
At both levels, we suggest that a sustainable solution requires a large initial investment to reach a critical mass, thus helping to attract and retain future women in systems.

### Acknowledgements {-}

We thank Kelly Shaw for her many insightful comments and suggestions.

### Disclaimer {-}

The opinions expressed in this viewpoint are the authors' own and do not reflect the views of Reed College or their employers.

<!-------------------------------------------------------------------------------------------------------------->

